---
title: "HP functions"
author: "So Yeon Park"
date: "2023-06-19"
output: pdf_document
---

1. Write a function using 4 inputs (lambda, alpha, beta, T) that returns event times; T is the end-time.
```{r}
hp_sim <- function(lambda, alpha, beta, T) {
  # empty vector for event times
  event_times <- numeric(0)
  
  # Set initial time = 0
  t <- 0
  
  # while loop
  while (t < T) {
    # calculate intensity at current time
    intensity <- lambda + sum(alpha * beta * exp(-beta * (t - event_times)))
    
    # generate the following event time using the intensity 
    event_time_next <- rexp(1, rate = intensity)
    
    # update current time
    t <- t + event_time_next
    
    # record new event time to the vector
    event_times <- c(event_times, t)
  }
  
  # return the simulated event times
  return(event_times)
}

# simulate using the following parameters
lambda <- 0.8
alpha <- 0.3
beta <- 1.4
T <- 10

# Simulate the Hawkes process
hp_events <- hp_sim(lambda, alpha, beta, T)

# Print the simulated event times
print(hp_events)

```

2. Modify the function by using "Ogata's modified thinning algorithm"
```{r}
hp_sim_omta <- function(lambda, alpha, beta, T) {
  # create an empty vector for event times
  event_times <- numeric(0)
  
  # set the initial time = 0
  t <- 0
  
  while (t < T) {
    # calculate the max intensity (calculated at the initial time t)
    intensity_max <- lambda + alpha * sum(exp(-beta * (t - event_times)))
    
    # generate the next event time with maximum intensity
    event_time_next <- rexp(1, rate = intensity_max)
    
    # generate a random number from uniform dist.
    u <- runif(1)
    
    # calculate current intensity at the next event time
    intensity_next <- lambda + alpha * sum(exp(-beta * (event_time_next - event_times)))
    
    # find the envelope/blanket function using Ogata's modified thinning algorithm
    envelope_f <- intensity_next / intensity_max
    
    # accept or reject?
    if (u <= envelope_f) {
      # accept the event
      t <- t + event_time_next
      
      # record new event time to the vector
      event_times <- c(event_times, t)
    }
  }
  
  # return simulated event times
  return(event_times)
}

# Simulate the Hawkes process
hp_omta_events <- hp_sim_omta(lambda, alpha, beta, T)

# Print the simulated event times
print(hp_omta_events)
```

3. Update with alpha < beta
```{r}
hp_sim_omta <- function(lambda, alpha, beta, T) {
  # for stationarity, I need alpha < beta
    if (alpha >= beta) {
    stop("Stationary hawkes process requires alpha<beta.")
} 
  # create an empty vector for event times
  event_times <- numeric(0)
  
  # set the initial time = 0
  t <- 0

  while (t < T) {
    # calculate the max intensity (calculated at the initial time t)
    intensity_max <- lambda + alpha * sum(exp(-beta * (t - event_times)))
    
    # generate the next event time with maximum intensity
    event_time_next <- rexp(1, rate = intensity_max)
    
    # generate a random number from uniform dist.
    u <- runif(1)
    
    # calculate current intensity at the next event time
    intensity_next <- lambda + alpha * sum(exp(-beta * (event_time_next - event_times)))
    
    # find the envelope/blanket function using Ogata's modified thinning algorithm
    envelope_f <- intensity_next / intensity_max
    
    # accept or reject?
    if (u <= envelope_f) {
      # accept the event
      t <- t + event_time_next
      
      # record new event time to the vector
      event_times <- c(event_times, t)
    }
  }
  
  # return simulated event times
  return(event_times)
}

# try with alpha >= beta
hp_sim_omta(lambda,alpha=0.8,beta=0.7,T)
hp_sim_omta(lambda,alpha=0.8,beta=0.8,T)
```